Supplement to Initial estuarine response to inorganic nutrient inputs from a legacy mining facility to Tampa Bay, Florida

### 2022-03-13 revision2/review_response.Rmd

This is a nice paper that addresses the effects of a high visibility/publicity wastewater spill on an iconic estuary (Tampa bay). Most of my comments were addressed, but there are still a few issues that need additional attention.

* __Response__: Thank you again for your comments. We have made efforts to address your final concerns as noted below.

1. The authors didn't address my question of how the plume model was verified, other than to point to a couple of papers. I read those papers and don't fully understand them as I am not a physical oceanographer, but as far as I could tell, neither paper clearly stated how verification was made using field observations. Given that this paper is likely to be read mostly by water quality experts and not physical oceanographers, I feel strongly that the authors need to include a brief (plain language… not physical oceanographer-speak) description of modeled plume verification, especially considering that the field sampling and interpretation of results is so strongly dependent on accurate quantification of plume evolution.

     * __Response__: The Tampa Bay Coastal Ocean Model was previously tested for sea level and velocity field veracity under extreme conditions, e.g., Hurricane Irma [@Chen18] and under more regular variations in tide, wind and river forcing by @Zhu15 and @Chen19.  Model results closely followed observations over the time scale associated with tides, wind forcing and the longer term averaged net estuarine circulation.  It was with this backdrop that we felt justified in adding a tracer component to examine how the Piney Point effluent would spread throughout Tampa Bay.  Two additional tests were then performed.  The first was to demonstrate that the model faithfully conserved the amount of tracer.  Thus we plotted what was known to be released and compared this with the integrated amount of tracer within the Tampa Bay Coastal Ocean model simulation.  Both were identical.  Next we compared patterns of modeled tracer distribution with ocean color observations by satellite remote sensing.  Again, the agreements were very good.  In essence, if a model is known to faithfully reproduce the three-dimensional circulation features, then it should perform similarly in tracking a passive scalar because the scalar is transported and spread by the circulation.  A dedicated paper of the modeling work (including model/data comparisons) is still in preparation. More details will be reported in that paper.  Preliminary results were reported at a recent conference (see Weisberg 2022, [BASIS7](https://tbep.org/estuary/basis/basis7/)).

### 2022-03-13 revision2/review_response.Rmd

This revised version is significantly improved compared to the first submission. However, there are still some issues that need addressing. It remains an overall descriptive study with little or no mechanistic understanding of why the bloom happened- but perhaps that is OK. The writing is better.

* __Response__: Thank you for your comments and we are pleased to hear that the draft has improved. We have addressed the remaining concerns below.

Title- no one other than local people will know what a Piney Point is. Needs revision to something along the lines of: Initial estuarine response to a spill of phosphate mining waste: time course of a red tide following  the Piney Point spill in Tampa Bay, USA

* __Response__: Agreed, the title has been changed to "Initial estuarine response to inorganic nutrient inputs from a legacy mining facility to Tampa Bay, Florida".  However, we have added "Piney Point" to the key words to assist with article indexing.

@@

This is a nice paper that addresses the effects of a high visibility/publicity wastewater spill on an iconic estuary (Tampa bay). Most of my comments were addressed, but there are still a few issues that need additional attention.

* __Response__: Thank you again for your comments. We have made efforts to address your final concerns as noted below.

1. The authors didn't address my question of how the plume model was verified, other than to point to a couple of papers. I read those papers and don't fully understand them as I am not a physical oceanographer, but as far as I could tell, neither paper clearly stated how verification was made using field observations. Given that this paper is likely to be read mostly by water quality experts and not physical oceanographers, I feel strongly that the authors need to include a brief (plain language… not physical oceanographer-speak) description of modeled plume verification, especially considering that the field sampling and interpretation of results is so strongly dependent on accurate quantification of plume evolution.

     * __Response__: The Tampa Bay Coastal Ocean Model was previously tested for sea level and velocity field veracity under extreme conditions, e.g., Hurricane Irma [@Chen18] and under more regular variations in tide, wind and river forcing by @Zhu15 and @Chen19.  Model results closely followed observations over the time scale associated with tides, wind forcing and the longer term averaged net estuarine circulation.  It was with this backdrop that we felt justified in adding a tracer component to examine how the Piney Point effluent would spread throughout Tampa Bay.  Two additional tests were then performed.  The first was to demonstrate that the model faithfully conserved the amount of tracer.  Thus, we plotted what was known to be released and compared this with the integrated amount of tracer within the Tampa Bay Coastal Ocean model simulation.  Both were identical.  Next, we compared patterns of modeled tracer distribution with ocean color observations by satellite remote sensing.  Again, the agreements were very good.  In essence, if a model is known to faithfully reproduce the three-dimensional circulation features, then it should perform similarly in tracking a passive scalar because the scalar is transported and spread by the circulation.  A dedicated paper of the modeling work (including model/data comparisons) is still in preparation. More details will be reported in that paper.  Preliminary results were reported at a recent conference (see Weisberg 2022, [BASIS7](https://tbep.org/estuary/basis/basis7/)).
